Interisland movements and oceanic swimming of woodland caribou in Newfoundland

### 2020-02-22 paper/manuscript.Rmd
fiddly refs

---
title: Interisland movements and oceanic swimming of woodland caribou in Newfoundland
author: Quinn M.R. Webber^1\*^, Jack G. Hendrix^1^, Alec L. Robitaille^2^,  Eric Vander Wal^1,2^
date: 
   ^1^Cognitive and Behavioural Ecology Interdisciplinary Program, Memorial University of Newfoundland
   ^2^Department of Biology, Memorial University of Newfoundland, Canada
   #\*Corresponding author: Quinn M.R. Webber (webber.quinn@gmail.com)
output: 
    github_document
  # bookdown::pdf_document2:
  #    number_sections: true
  #    toc: false
  # bookdown::word_document2:
    #reference_docx: template.docx
bibliography: ../references/swim.bib
#csl: csl.csl
---

@@

# Summary and conclusion
For several decades after their introduction to Fogo Island, the caribou
population did not exceed ~100 individuals [@Bergerud1989 ; Newfoundland and
Labrador Wildlife Division, unpublished data). During the 1990s, population
density reached 300 animals and anecdotal evidence suggests that [only then???]
caribou began to occupy other islands in the archipelago. We surmise that as
competition among conspecifics increased along with population size,
density-dependent habitat selection resulted in expansion of the population to
nearby islands. Following Ideal Free Distribution theory, caribou should swim to
new islands when the average fitness of individuals on the starting island
exceeds the density-fitness equilibrium [@Morris1987]. While this is an ultimate
explanation to a series of proximate observations, fine-scale density-dependent
habitat selection is theoretically possible [@Webber2018]. Predator avoidance
could also contribute to caribou swimming behaviour, but given the presumed low
density of coyotes on Fogo Island and that predation primarily only affects
young calves, we suspect predator is a possible, but less likely explanation.
<!-- this last bit is sort of me bullshitting, not sure it's actually defensible? 
QW: changed it a bit -->

>>>>

---
title: Interisland movements and oceanic swimming of woodland caribou in Newfoundland
author: Quinn M.R. Webber^1\*^, Jack G. Hendrix^1^, Alec L. Robitaille^2^,  Eric Vander Wal^1,2^
date: 
   ^1^Cognitive and Behavioural Ecology Interdisciplinary Program, Memorial University of Newfoundland
   ^2^Department of Biology, Memorial University of Newfoundland, Canada
   #\*Corresponding author: Quinn M.R. Webber (webber.quinn@gmail.com)
output: 
   # github_document
  # bookdown::pdf_document2:
  #    number_sections: true
  #    toc: false
   bookdown::word_document2:
    reference_docx: template.docx
bibliography: ../references/swim.bib
#csl: csl.csl
---

@@

# Summary and conclusion
For several decades after their introduction to Fogo Island, the caribou
population did not exceed ~100 individuals [@Bergerud1989, Newfoundland and
Labrador Wildlife Division, unpublished data]. During the 1990s, population
density reached 300 animals and anecdotal evidence suggests that [only then???]
caribou began to occupy other islands in the archipelago. We surmise that as
competition among conspecifics increased along with population size,
density-dependent habitat selection resulted in expansion of the population to
nearby islands. Following Ideal Free Distribution theory, caribou should swim to
new islands when the average fitness of individuals on the starting island
exceeds the density-fitness equilibrium [@Morris1987]. While this is an ultimate
explanation to a series of proximate observations, fine-scale density-dependent
habitat selection is theoretically possible [@Webber2018]. Predator avoidance
could also contribute to caribou swimming behaviour, but given the presumed low
density of coyotes on Fogo Island and that predation primarily only affects
young calves, we suspect predator risk is a possible, but less likely
explanation for swimming.
<!-- this last bit is sort of me bullshitting, not sure it's actually defensible? 
QW: changed it a bit - how is this? -->
